A Simplified Formulation for Molecular Dynamics with Rigid Bodies

### 2017-01-31 rigid_bodies.tex
nve

\begin{equation}
\label{eq:H_split_omega}
\mathcal{H} = \frac{\tr{\vt p} \vt p}{2m} + U(\vt r, \vt q) + \frac{1}{2} \sum_{k=1}^3 I_k \omega_k^2.
\end{equation}

where an index $i$ ranging from $1$ to $N$ indicates the quantities related to each body, with $m_i$ is the mass and  $\mt I_i$ the inertia tensor which in the principal frame is a diagonal matrix withe entries given by ($I_1$,$I_2$,$I_3$). The equations of motion 
\begin{subequations}
\label{eq:EDO_system}
\begin{align}
&\dot{\vt r} = \frac{1}{m} \vt p \\
&\dot{\vt p} = \mt F \\
&\dot{\vt q} = \frac{1}{2} \sum_{k=1}^3 \omega_k \hat{\mt B}_k \vt q \label{eq:edo_q} \\
&\dot{\vt \pi} = \frac{1}{2} \sum_{k=1}^3 \omega_k \hat{\mt B}_k \vt \pi + 2 \mt C \vt \tau \label{eq:edo_pi}
\end{align}
\end{subequations}

consider the entries of $\vt \omega$ individually. Given that $2{\mt I}{\vt \omega} = \tr{\mt B}{\vt \pi}$, according to Eq.~\ref{eq:vector_entries}, each entry of $\vt \omega$ is obtained by
\begin{equation}
\label{eq:omega_entry}
\omega_k = \frac{\tr{\vt \pi} \hat{\mt B}_k \vt q}{2 I_k},
\end{equation}
where $\hat{\mt B}_k$ is one of the permutation matrices presented in Appendix~\ref{sec:auxiliary_math} of the aforementione peper.

%Finally, we open the products $\mt B \vt \omega$ and $\mt \Omega \vt \omega$ as in Eq.~\ref{eq:product_B_vector}, which leads us to the following new form for the equations of motion in Eqs.~\ref{eq:EDO_q} and \ref{eq:EDO_pi}:
%\begin{subequations}
%\label{eq:EDO_system_alternative}
%\begin{align}
%&\dot{\vt q} = \frac{1}{2} \sum_{k=1}^3 \omega_k \hat{\mt B}_k \vt q \label{eq:edo_q} \\
%&\dot{\vt \pi} = \frac{1}{2} \sum_{k=1}^3 \omega_k \hat{\mt B}_k \vt \pi + 2 \mt C \vt \tau %\label{eq:edo_pi}
%\end{align}
%\end{subequations}

>>>>

The equations of motion are obtained from the Hamiltonian by $\dot{\vt r} = \partial \mathcal{H} / \partial \vt p$, $\dot{\vt p} = -\partial \mathcal{H} / \partial \vt r$, $\dot{\vt q} = \partial \mathcal{H} / \partial \vt \pi$, and $\dot{\vt \pi} = -\partial \mathcal{H} / \partial \vt q$.\cite{Goldstein2002}. some algebra manipulations where we defined $\mt \Omega = \mt B(\vt \pi)$, it follows that $\tr{\mt B}{\vt \pi} = - \tr{\mt \Omega}{\vt q}$. For a rigid body composed of $n_p$ individual particles, the derivative $\partial U/\partial \vt r$ is obtained by
\[
\diff{U}{\vt r} = -\vt F = -\sum_{j=1}^{n_p} {\vt F_j},
\]
where $\vt F_j$ is the force applied on particle $j$, expressed in the space-fixed frame of reference and thus $\vt F$ is the resultant force on the body. As demonstrated in Abreu, we have
\[
\diff{U}{\vt q} = -2 \mt C \vt \tau = -2 \mt C \sum_{j=1}^{n_p} {\vt \delta_j} \times {\vt F_j},
\]
where $\vt \tau$ is the resultant torque exerted on the body, also expressed in the space-fixed frame. e. Finally, after carrying out the required differentiations of $\mathcal{H}$  and substituting $\tr{\mt B} \vt \pi = - \tr{\mt \Omega} \vt q = 2 {\mt I} \vt \omega$, we obtain the following Hamiltonian system of equations of motion for a rigid body:
\begin{subequations}
\label{eq:EDO_system}
\begin{align}
&\dot{\vt r} = \frac{1}{m} \vt p \\
&\dot{\vt p} = \mt F \\
&\dot{\vt q} = \frac{1}{2} \mt B \vt \omega \label{eq:EDO_q} \\
&\dot{\vt \pi} = \frac{1}{2} \mt \Omega \vt \omega + 2 \mt C \vt \tau \label{eq:EDO_pi}
\end{align}
\end{subequations}

we now consider the entries of $\vt \omega$ individually. Given that $2{\mt I}{\vt \omega} = \tr{\mt B}{\vt \pi}$, according to Eq.~\eqref{eq:vector_entries}, each entry of $\vt \omega$ is obtained by
\begin{equation}
\label{eq:omega_entry}
\omega_k = \frac{\tr{\vt \pi} \hat{\mt B}_k \vt q}{2 I_k},
\end{equation}
where $\hat{\mt B}_k$ is one of the permutation matrices presented in Appendix~\ref{sec:auxiliary_math}. In this context, the Hamiltonian of the rigid body can be rewritten as
\begin{equation}
\label{eq:H_split_omega}
\mathcal{H} = \frac{\tr{\vt p} \vt p}{2m} + U(\vt r, \vt q) + \frac{1}{2} \sum_{k=1}^3 I_k \omega_k^2.
\end{equation}

Finally, we open the products $\mt B \vt \omega$ and $\mt \Omega \vt \omega$ as in Eq.~\eqref{eq:product_B_vector}, which leads us to the following new form for the equations of motion in Eqs.~\eqref{eq:EDO_q} and \eqref{eq:EDO_pi}:
\begin{subequations}
\label{eq:EDO_system_alternative}
\begin{align}
&\dot{\vt q} = \frac{1}{2} \left( \sum_{k=1}^3 \omega_k \hat{\mt B}_k \right) \vt q \label{eq:edo_q} \\
&\dot{\vt \pi} = \frac{1}{2} \left( \sum_{k=1}^3 \omega_k \hat{\mt B}_k \right) \vt \pi + 2 \mt C \vt \tau \label{eq:edo_pi}
\end{align}
\end{subequations}

where an index $i$ ranging from $1$ to $N$ indicates the quantities related to each body, with $m_i$ is the mass and  $\mt I_i$ the inertia tensor which in the principal frame is a diagonal matrix withe entries given by ($I_1$,$I_2$,$I_3$).
